---
title: "Problem set 3 - part 2"
author: "Alin Ierima"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

# Ideology of US Supreme Court Justices

We learned an important programming method called the *for loop*. In
this part of the exam you will practice using loops with data on the
ideological positions of United States Supreme Court Justices. Just like
legislators, justices make voting decisions that we can use to estimate
their ideological positions.

This exercise is based in part on Andrew Martin and Kevin Quinn. (2002).
'Dynamic Ideal Point Estimation via Markov Chain Monte Carlo for the
U.S. Supreme Court, 1953-1999.' *Political Analysis*, 10:2, pp.134-154.

The file `justices.csv` contains the following variables:

| Name      | Description                                   |
|:----------|:----------------------------------------------|
| `term`    | Supreme Court term (a year)                   |
| `justice` | Justice's name                                |
| `idealpt` | Justice's estimated ideal point in that term  |
| `pparty`  | Political party of the president in that term |
| `pres`    | President's name                              |

The ideal points of the justices are negative to indicate liberal
preferences and positive to indicate conservative preferences.

## Question 1 (10 points)

We would like to know the *median* ideal point for the Court during each
term included in the dataset.

-   First, calculate the median ideal point (variable `idealpt`) for
    each term (variable `term`) of the Court. (4 points)
-   Next, generate a plot with `term` on the horizontal axis (x-axis)
    and the median ideal points on the vertical axis (y-axis). (3
    points)
-   Connect all points with a line (You can find good examples how to do
    so here: <https://r-coder.com/plot-r/#R_plot_type>) (1 point)
-   Include a dashed horizontal line at zero to indicate a "neutral"
    ideal point. (1 point)
-   Be sure to include informative axis labels and a plot title. (1
    point)

## Answer 1

```{r}
justices <- read.csv("data/justices.csv", stringsAsFactors = FALSE)
```

```{r}
median_idealpt=tapply(justices$idealpt, justices$term, median)
justices$median_idealpt = median_idealpt[as.character(justices$term)]
```

```{r}
plot(unique(justices$term), median_idealpt, xlab="Terms", ylab="Median ideal points", type = "o",pch = 7, ylim = c(-1.0, 1.3), main="The median ideal points of Justices across the years" )
abline(h=0, lty=3, col="steelblue4")
text(1991, -0.1, "Neutral ideal point")
```

## Question 2 (20 points)

Next, we would like to identify the name of the justice with an ideal
position equal to the median ideal point **for each term**. Use a for
loop to do so!

The loop should: (14 points)

-   go though each term
-   identify the justice whose ideal position is equal to the the median
    position in this term
-   save the names of the justices for each term in a container object
    called `median_justice`.

In addition, answer in the text the following questions. [Instead of
copy-pasting numbers in your answer, please use 'r some code'. For
example, if you want to refer to the mean value of the ideal positions
of all justices, instead of writing, the mean ideal position is "
-0.008321182", write the mean ideal position is
`r mean(justices$idealpt)`, see how this looks in Rmarkdown.]

1)  Which justice had the median ideal point in the most terms? (3
    points)
2)  How long did this justice serve on the Court overall? (2 points)
3)  What was this justice's average ideal point over his/her entire
    tenure on the Court? (1 points)

## Answer 2

**Median justices by term**

```{r}
tapply(justices$justice, justices$term, median)
```

**Number of times a given justice had a position equal to the median
term position**

```{r}
median_justice = NA
for (i in unique(justices$term)) {
  justice_name= justices$justice[justices$term == i & justices$idealpt == median_idealpt[as.character(i) ]]
  median_justice=c(median_justice, ifelse(length(justice_name) > 0,justice_name, NA) )
}

justice_counts = table(median_justice)
justice_counts
```

**How long did this justice serve on the Court overall?**

```{r}
length(justices$term[justices$justice == "White"])
```

**What was this justice's average ideal point over his/her entire tenure
on the Court?**

```{r}
mean(justices$idealpt[justices$justice== "White"])
```

[Insert your answer here:]{style="color: green;"} In this exercise we
wanted to identify the justice with an ideal position equal to the
median point, therefore we took White since he he has the biggest length
of serving, `32`, and his average ideal point over the entire tenure was
`0.4401563.`

## Question 3 (30 points)

We now turn to the relationship between Supreme Court ideology and the
president. Specifically, we want to see how the ideology of the Supreme
Court changes over the course of each president's time in office.

-   Begin by creating two empty 'container' vectors: one to hold
    Democratic presidents (name it `dem_pres_shift`), and another for
    Republican presidents( (name it `rep_pres_shift`). Label each vector
    with the presidents' names. (6 points)
-   Next, using a loop, calculate for each Democratic president the
    shift in Supreme Court ideology by subtracting the Court's median
    ideal point in the president's first term from its median ideal
    point in the president's last term. (10 points)
-   Store these values in your Democratic container vector
    (`dem_pres_shift`). (2 points)
-   Repeat the same process for Republican presidents. (Hint: you can
    create a second loop.) (12 points)

## Answer 3

**Empty container `dem_pres_shift`**

```{r}
dem_pres_shift = rep(NA,length(unique(justices$pres[justices$pparty =="D"])))
names(dem_pres_shift) =unique(justices$pres[justices$pparty == "D"])
dem_pres_shift
```

**Empty container `rep_pres_shift`**

```{r}
rep_pres_shift = rep(NA,length(unique(justices$pres[justices$pparty =="R"])))
names(rep_pres_shift) = unique(justices$pres[justices$pparty == "R"])
rep_pres_shift
```

**Ideological change in the Supreme Court during Democratic presidents**

```{r}


for (i in names(dem_pres_shift)) {
  dem_pres_data = subset(justices, justices$pres == i &pparty == "D")
  
  first_term = min(dem_pres_data$term, na.rm=T)
  last_term = max(dem_pres_data$term, na.rm=T)
  
   if (is.na(first_term) | is.na(last_term) ) {
  shift=NULL   
   } else{
       median_first = median(dem_pres_data$idealpt[as.numeric(dem_pres_data$term) == first_term], na.rm=T)
  median_last = median(dem_pres_data$idealpt[as.numeric(dem_pres_data$term) == last_term],na.rm=T)
       shift = median_last - median_first
   }              

  dem_pres_shift[i] = shift
}
  dem_pres_shift

```

**Ideological change in the Supreme Court during Republican presidents**

```{r}
for (i in names(rep_pres_shift)) {
  rep_pres_data = subset(justices, justices$pres == i &pparty == "R")
  first_term = min(rep_pres_data$term, na.rm=T)
  last_term = max(rep_pres_data$term, na.rm=T)
   if (is.na(first_term) | is.na(last_term) ) {
  shift=NULL   
   } else{
       median_first = median(rep_pres_data$idealpt[as.numeric(rep_pres_data$term) == first_term], na.rm=T)
  median_last = median(rep_pres_data$idealpt[as.numeric(rep_pres_data$term) == last_term],na.rm=T)
     
   }              

  shift = median_last - median_first
  rep_pres_shift[i] = shift
}
  rep_pres_shift
```

## Question 4 (10 points)

-   What was the mean and standard deviation of the Supreme Court
    ideology shifts you just calculated when looking only at the
    Democratic presidencies?
-   What about the Republican presidencies?
-   Which Republican president's tenure had the largest conservative
    (positive) shift on the Court?
-   Which Democratic president's tenure had the largest liberal
    (negative) shift?

## Answer 4

**Mean and standard deviation of the Supreme Court ideology shifts
(Democratic presidencies)**

```{r}
mean(dem_pres_shift)
sd(dem_pres_shift)
```

**Mean and standard deviation of the Supreme Court ideology shifts
(Republican presidencies)\\**

```{r}
mean(rep_pres_shift)
sd(rep_pres_shift)
```

**Democratic president with the largest conservative (positive) shift in
the Court**

```{r}
names(dem_pres_shift[dem_pres_shift == min(dem_pres_shift)])
min(dem_pres_shift)
```

**Republican president's with the largest liberal (negative) shift in
the Court?**

```{r}
names(rep_pres_shift[rep_pres_shift == max(rep_pres_shift)])
max(rep_pres_shift)
```

## Question 5 (30 points)

In this task you will create a plot which shows the median Supreme Court
ideal point over time together with the ideological development of each
justice over time.

A similar plot was recently features in the Economist after the death of
Ruth Bader Ginsburg in 2020. Those interested in the article can read it
here:
<https://www.economist.com/graphic-detail/2020/09/21/the-supreme-court-may-be-about-to-take-a-hard-right-turn>
. CEU's library has a free access to articles from `The Economist`.

Do the following steps:

-   Create a plot that shows the median Supreme Court ideal point over
    time. Use a thick black line (thickness equal to 4). (3 points) [The
    figure below uses transparency level equal to 0.7]
-   Then, add lines for the ideal points of each unique justice to the
    same plot. Use a `for loop` for this purpose. (10 points)
-   The color of each line should be `orangered` if the justice was
    appointed by a Republican and `steelblue4` if he or she was
    appointed by a Democrat. (You can assume that when a Justice first
    appears in the data, they were appointed by the president sitting
    during that term.) (10 points)
-   Label each line with the justice's last name. (5 points)
-   Briefly discuss what trends in the ideology of justices you see over
    time. (2 points)

## Answer 5

Insert your plot here:

```{r}

plot(justices$term, justices$median_idealpt, type= "l", lwd=4, ylim=c(-10,10), xlim=c(1940,2010), ylab = "Justices' ideology", xlab="Year", main="Supreme Court Ideological Evolution")
  for (i in unique(justices$justice)) {
    lines(x = justices$term[justices$justice == i], y = justices$idealpt[justices$justice == i], col=col )
    if (justices$pparty[justices$justice == i][1]=="R") {
      col = "orangered" 
    }
      else { col = "steelblue4"}
    indices = justices$justice == i
    last_term = tail(justices$term[indices], 1)
  last_idealpt = tail(justices$idealpt[indices], 1)
 text(x = last_term +0.5,
       y = last_idealpt-0.5,
       labels = i,
       col = col, cex = 0.5)
}
abline(h=0)

```

[Insert your answer here:]{style="color: green;"} It seems to me that
the general trend is towards more liberal/progressive republican judges,
which is quite interesting. According to the median points, not much h
as changed in the last fifty years, however through this graph we can
see that there are quite a lot of republican justices that slided
further to the liberal ideology, some being more democrat than the
democrats themselves.Also in the 40s and 50s, it seems the norm was for
judges to be more conservative (including most of the democrat justices)

## Question 6 (Bonus question for 20 points)

Your boss really likes the graph you created in question 5, great job!
She thinks that the graph can be tailored a bit. In, particular, she
asks you to do the following changes:

1)  Make the ideal points lines of the last justices - those who were
    still justices between 2000 and 2008 - thicker. (5 points)
2)  Add a point for the starting ideological point of each justice. Make
    these points for the last justices bigger than for the remaining
    justices. (5 points)
3)  Make the names of the last justices readable (change their
    positions, so that the names do not overlap) and color them
    according to the party color of the president that appointed these
    justices. (5 points)
4)  Include the following information: (5 points)

-   Label the conservative and liberal side of the ideological spectrum
    using arrows
-   Add a label for the median justice line.
-   Add a legend for the colors of the justices ideal points. Make the
    same legend as in the example below. Notice that this legend has no
    box, the legend points are horizontal and there is a title.

Hints:

-   To specify which justices' should have thicker lines and bigger
    points, you can use different conditional statements within the loop
    (e.g. `if(){}, if(){}else{}` etc.).
-   If you would like to find out which justices were justices between
    2000 and 2008 you can use the value matching `%in%` operator. This
    operator allows you to compare which elements in a given vector
    (e.g. vector x) occur in another vector (e.g. vector y). For
    example, let's say you have a vector x with the numbers
    `c(1,3,2,5,4)` and another vector y with the numbers `c(3,4,6)`. You
    would like to know which elements in x have a match in y. To do so
    you can run `x %in% y` and R will return a vector with TRUE and
    FALSE for each element in x. TRUE indicates that a given element in
    x had a match in y.

```{r}
x <- c(1,3,2,5,4)
y <- c(3,4,6)

res_match <- x %in% y
res_match
# You can count how many elements were matched by using `sum(res_match)`
sum(res_match)
```

-   remember that you can adjust the colors to be transparent by
    applying the function
    `adjustcolor("colorname", number_for_transparency_level)`. The graph
    below uses most of the time `adjustcolor("orangered", 0.8)` and
    `adjustcolor("steelblue4", 0.8)`

-   you can draw arrows using the function `arrows()`, where you need to
    specify the coordinates of the arrow using the arguments `x0`, `y0`
    (starting point of the arrow on the x and y coordinates
    respectively), `x1`, `y1` (end point of the arrow on the x and y
    coordinates respectively). Additionally, the following arguments
    were specified in the `arrows()` function used in the below plot:
    `length = 0.05`, `angle = 30`, `code = 2`, `col = steelblue4`.

-   To learn more how to draw a legend with specific features, type
    `?legend` in the console. The R documentation provides detailed
    information on many features that can be specified in the legend.
    For example, you find information about the type of the box to be
    drawn around the legend (argument `bty`), whether to set the legend
    horizontal or vertical (argument `horiz`), the color, adjustment and
    the text of the title within the legend (arguments `title.col`,
    `title.adj`, `title`)

## Answer 6

Insert your plot here:

```{r}
plot(justices$term, justices$median_idealpt, type= "l", lwd=4, ylim=c(-8,8), xlim=c(1940,2010), ylab = "Justices' ideology", xlab="Year", main="Supreme Court Ideological Evolution")
  for (i in unique(justices$justice)) {
    lines(x = justices$term[justices$justice == i], y = justices$idealpt[justices$justice == i], col=col )
    if (justices$pparty[justices$justice == i][1]=="R") {
      col = "orangered" 
    }
      else { col = "steelblue4"}
    indices = justices$justice == i
    last_term = tail(justices$term[indices], 1)
  last_idealpt = tail(justices$idealpt[indices], 1)
 text(x = last_term +1,
       y = last_idealpt-0.5,
       labels = i,
       col = col, cex = 0.4)
 
     if (tail(justices$term[justices$justice == i],1) >= 2000 & tail(justices$term[justices$justice == i],1)<= 2008) {
      lines(x = justices$term[justices$justice == i], y = justices$idealpt[justices$justice == i], col = col, lwd = 2, cex=1)
    } else {
    }
}
abline(h=0)
arrows(x0 = 1940, y0 = -5, y1 = -7, length = 0.05, angle = 30, code = 2, col = "steelblue4")
text(x=1945, y=6, "More \nConservative", cex= 0.6, col="steelblue4")
arrows(x0=1940, y0=5, y1 = 7, length=0.05, angle=30, code=2, col="steelblue4")
text(x=1943, y=-6, "More \nLiberal", cex= 0.6, col="steelblue4")
legend("bottomright",legend = c("Democrat", "Republican"), fill = c("steelblue4", "orangered"), title = "Appointed by a", bty = "n", text.col = c("steelblue4", "orangered"),horiz=T, cex = 0.7)


```

## Evaluation

-   5 questions for a total of 100 points
-   1 bonus question for a total of 20 points
