---
title: "Data Report 1"
author: "Alin Ierima"
output:
  pdf_document: default
  html: default
---

# Dataset Description

According to the documentation, `Palmerpenguins` is a database that includes information different characteristics of adult penguins near the Palmer Station, Antarctica. We will take a look at the `penguins` dataset, which includes the following variables: species, the island, bill length and depth (mm), flipper length (mm), body mass (gr), sex and year.

```{r installing/library}
#| eval: false
#| include: false
install.packages("palmerpenguins") 
library("palmerpenguins")
```

# Cleaning

As discussed in class, the data is already considered clean.

# Exploration

The species of the penguins seem to have a great importance for their length. As it can be seen in figure 1, **Adelie penguins are overall the smallest** in terms of size, **followed by Chinstrap** (bigger bills and slightly bigger flippers than the previous) **and, lastly, Gentoo** (bigger in both variables).

```{r Figure 1, warning=FALSE, echo=FALSE, message=FALSE}
library(tidyverse)
library(palmerpenguins)
ggplot(data = penguins, 
       mapping = aes( 
         penguins$bill_length_mm, #I know it is a little iterative but you cannot go wrong with $'s
         penguins$flipper_length_mm,
         color = species)) + geom_point() + labs(
           x = "The Bill Length of Penguins (milimeters)",
           y = "The Flipper Length of Penguins (milimeters)", 
           #title = "Size Differences Amongst Different Species", this makes the plot look ugly
         ) + theme(legend.background = element_rect(fill="gray70"))
 


```

Does the island penguins come from influence their size? **Penguins from Biscoe seem to be bigger than the rest,** and they mostly overlap with the Gentoo species as shown in the previous figure. The reason is a matter of speculation, as there might be a reason why this species thrive on the island.

```{r Figure 2, warning=FALSE, echo=FALSE, message=FALSE}
ggplot(data = penguins, 
       mapping = aes( 
         penguins$bill_length_mm,
         penguins$flipper_length_mm,
         color = island)) + geom_point() + labs(
           x = "The Bill Length of Penguins (milimeters)",
           y = "The Flipper Length of Penguins (milimeters)", 
         ) + theme(legend.background = element_rect(fill="gray70"))
 
```

Let's further delve into the composition of each island. Adelie penguins, which are the smallest, seem to be endemic to all three islands, whereas **Chinstrap and Gentoo remain geographically confined** to only one separate island. **Togersten is** **only populated** **by Adelie species**.

```{r Figure 3, warning=FALSE, echo=FALSE, message=FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(position = position_stack(), stat = "count") +
  labs(
    x = "Island",
    y = "Count",
    title = "Penguins of Each Species by Island"
  ) +
  theme_minimal()
```

Intuitively enough, Gentoo penguins are not only the largest but also the heaviest. However, Adelie and Chinstrap penguins seem to be quite similar in their heaviness, median-wise.

```{r Figure 4, warning=FALSE, echo=FALSE, message=FALSE}
library(tidyverse)
library(palmerpenguins)
ggplot(penguins, aes(species, body_mass_g, fill = species)) + geom_boxplot() +
  labs(
      x = "Species",
    y = "The body mass of Penguins (grams)"
  ) + theme(legend.position = "none")
 
```

At the same time, **sex** does seem to make a difference when it comes to length. Scatterplotting seems not to help a great deal in this case, but out of this figure we can take out that when it comes to flipper length, the outliers (those with more than 225mm) **are exclusively male**.

```{r Figure 5, warning=FALSE, echo=FALSE, message=FALSE}
library(tidyverse)
library(palmerpenguins)
ggplot(data = penguins, 
       mapping = aes( 
         penguins$bill_length_mm, 
         penguins$flipper_length_mm,
         color = sex)) + geom_point() + labs(
           x = "The Bill Length of Penguins (milimeters)",
           y = "The Flipper Length of Penguins (milimeters)", 
         ) + theme(legend.background = element_rect(fill="gray70"))
 
```

The following two boxplots clear the picture when it comes to sex and penguin dimensions. **Males are, on average, bigger than females** and also represent the maximum recorded values. NA's of sex are added for representation and transparency purposes, even though their number is quite small (11 compared to 333).

```{r counting NAs, warning=FALSE, echo=FALSE, message=FALSE, results = "hide"}

sum(is.na(penguins$sex)) # so there are eleven NA s 
sum(!is.na(penguins$sex)) #... and 333 complete entries
```

```{r Figure 6, warning=FALSE, echo=FALSE, message=FALSE}
library(tidyverse)
library(palmerpenguins)
ggplot(data = penguins, 
       mapping = aes( 
         penguins$sex,
         penguins$flipper_length_mm,
         fill = sex)) + geom_boxplot() + labs(
           x = "Sex of Penguins",
           y = "The Flipper Length of Penguins (milimeters)", 
         ) + theme(legend.background = element_rect(fill="gray70"))
 
```

```{r Figure 7, warning=FALSE, echo=FALSE, message=FALSE}
library(tidyverse)
library(palmerpenguins)
ggplot(data = penguins, 
       mapping = aes( 
         penguins$sex,
         penguins$bill_length_mm,
         fill = sex)) + geom_boxplot() + labs(
           x = "Sex of Penguins",
           y = "The Bill Length of Penguins (milimeters)", 
         ) + theme(legend.background = element_rect(fill="gray70"))
 
```

When it comes to body mass, **male penguins are also heavier** and, as shown in the two figures below, they represent the biggest penguins from the dataset.

```{r Figure 8, warning=FALSE, echo=FALSE, message=FALSE}
library(tidyverse)
library(palmerpenguins)
ggplot(data = penguins, 
       mapping = aes( 
         penguins$body_mass_g, 
        fill=sex)) + geom_histogram(binwidth = 80, alpha = 0.8) + labs(
           x = "Body Mass (grams)",
           y = "Frequency", 
         ) + theme(legend.background = element_rect(fill="gray70"))
```

```{r Figure 9, warning=FALSE, echo=FALSE, message=FALSE}
ggplot(penguins, aes(sex, body_mass_g, fill = sex)) + geom_boxplot() +
  labs(
    x = "Sex",
    y = "The body mass of Penguins (grams)"
  ) + theme(legend.position = "none")
```

# Conclusion (15 points)

To summarize the findings, the following factors seem to affect the mass and sizes of penguins:

-   Species: Gentoos are the biggest and heaviest, followed by Chinstrap and Adelie

-   Island, as the largest penguins live on Biscoe island; however, Adelie penguins live on all three islands. It is unclear whether only Biscoe offers the needed conditions for bigger penguins to live there

-   Sex: male penguins are bigger and heavier than female penguins

# Code exercises

Use the `msleep` data (included with `ggplot2`) for both exercises.

**Exercise:** Get all the animals who are heavier than the average bodyweight in the data; keep only the `name` and `bodywt` columns; and order the rows by `bodywt` in a descending order.

```{r Exercise 1}

avg_weight <- mean(msleep$bodywt, na.rm = T) #calculating the mean

msleep |> filter(bodywt > avg_weight) |> 
  select(name, bodywt) |> 
  arrange(desc(bodywt)) #filter by big animals, select only two columns, arrange desc by bodywt

```

**Exercise 5:** Start fresh (not with the data from exercise 4). Create a new column called `brainwt_ratio` showing the ratio of of brain mass to the total body weight. Round the ratio to 4 digits. Keep the `name` and `brainwt_ratio` colums and keep the 10 animals with the highest ratio.

```{r}
msleep |> 
  mutate(brainwt_ratio = round (msleep$brainwt / msleep$bodywt, 4)) |> 
  select(name, brainwt_ratio) |>
  arrange(desc(brainwt_ratio)) |>
  head(10) #only the first 10

```
